Brother, Can You Spare a Manufacturing Job? How Voters React to Deindustrialization

### 2021-12-12 Final Draft/Final Draft.Rmd
figure fix, small changes

\FloatBarrier
\begin{figure} \label{ManuMap}
\caption{Concentration of Deindustrialization (2004-2015) in the eastern United States }
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center" , regfloat=FALSE}
# load map
model_input <- readRDS("../model_5_input.rds")

@@

<!-- FIGURE X: TIME SERIES OF MANUFACUTRING EMPLOYMENT  -->
\FloatBarrier
\begin{figure} \label{natlPlotTS}
\caption{Nationwide Change in Manufacturing Employment, 1995-2019}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}
# manufacturing layoffs over time

@@

stargazer(regResultList$secondStageModelNetC04A,
          regResultList$secondStageModelNetC04,
          matchResultList$secondStageModelOurs04LW_3,
          regResultList$secondStageModelWNetC04A,
          regResultList$secondStageModelWNetC04,
          matchResultList$secondStageModelOurs04LW_6,
          se = list(
            regResultList$rseNetC04A_2,
            regResultList$rseNetC04_2,
            matchResultList$rseOurs04LW3_2,
            regResultList$rseWNetC04A_2,
            regResultList$rseWNetC04_2,
            matchResultList$rseOurs04LW6_2
          ),
          title = "Effect of Manufacturing Job Loss from 2004 to 2015",
          dep.var.labels = "Change in Dem. Share, 2016-2012",
          omit.stat = c("rsq","ser"),
          keep = c("Job Losses", "White Job Losses", "Nonwhite Job Losses"),
          add.lines = list(
            c("White Share/Svc Layoffs", "No", "Yes", "Yes", "No", "Yes", "Yes"),
            c("Covariate Balancing", "No", "No", "Yes", "No", "No", "Yes")
          ),
          digits = 2,
          header = F,
          type = "latex",
          label = "regResult04"
)

@@

\FloatBarrier
\begin{figure} \label{counterfactual1}
\caption{Counterfactual Scenario: Change in Democratic Vote Share Assuming 25th Percentile of Manufacturing Job Losses}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}

@@

\FloatBarrier
\begin{figure} \label{counterfactual2}

\caption{Counterfactual Simulation Results: Democratic Electoral College Votes Assuming White Manufacturing Job Losses Between 1st and 99th Percentile }
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}

@@

\FloatBarrier
\begin{figure} \label{counterfactual3}
\caption{Counterfactual Simulation Results: Democratic Presidential Win Frequency}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}

@@

Restricting our analysis to only the 2012-2015 period, while continuing to use the net job loss metric, produces an effect opposite to that of \cite{Baccini21}. The loss of Democratic vote share seems to be associated with a gain in manufacturing jobs, both in general and among whites. We believe this time-restricted analysis suffers from an omitted variable bias. As discussed above, the 2012-2015 period can be broadly characterized as a slight rebound from the manufacturing job losses in the 2008 financial crisis. The counties that experience job losses from 2004-2012 are more likely to have job gains in 2012-2015 (see \ref{vShapePlot}). An apparent relationship between 2012-2015 job gains and loss of Democratic vote share in fact represents 2004-2012 job losses that may lead to both the later job gains and the loss of vote share. These results can be seen in Table \ref{regResult12}.
\FloatBarrier
```{r,  echo=F, message = F, results = "asis", regfloat=FALSE}

@@

\begin{figure} \label{vShapePlot1}
\caption{V-Shaped Recovery in Manufacturing Job Losses: Manufacturing Job Losses (2004-2011) by (2004-2011) Decile Groupings}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center"}

@@

\begin{figure} \label{vShapePlot2}
\caption{V-Shaped Recovery in Manufacturing Job Losses: Manufacturing Job Losses (2012-2015) by 2004-2011 Decile Groupings}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center"}

@@

\begin{figure} \label{balPlot}
\caption{Covariate Balance in Aggregate}

```{r, message = F, fig.width = 5.5, fig.align="center"}
cbpsObjs <- readRDS("../cbpsObjects.rds")
plot(cbpsObjs$fit04)

@@

\begin{figure} \label{balPlotWhite}
\caption{Covariate Balance in Whites}

```{r, message = F, fig.width = 5.5, fig.align="center"}
plot(cbpsObjs$fit04_white)

>>>>

\FloatBarrier
\begin{figure} 
\caption{Concentration of Deindustrialization (2004-2015) in the eastern United States }
\label{ManuMap}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center" , regfloat=FALSE}
# load map
model_input <- readRDS("../model_5_input.rds")

@@

<!-- FIGURE X: TIME SERIES OF MANUFACUTRING EMPLOYMENT  -->
\FloatBarrier
\begin{figure}
\caption{Nationwide Change in Manufacturing Employment, 1995-2019}
\label{natlPlotTS}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}
# manufacturing layoffs over time

@@

stargazer(regResultList$secondStageModelNetC04A,
          regResultList$secondStageModelNetC04,
          matchResultList$secondStageModelOurs04LW_3,
          regResultList$secondStageModelWNetC04A,
          regResultList$secondStageModelWNetC04,
          matchResultList$secondStageModelOurs04LW_6,
          se = list(
            regResultList$rseNetC04A_2,
            regResultList$rseNetC04_2,
            matchResultList$rseOurs04LW3_2,
            regResultList$rseWNetC04A_2,
            regResultList$rseWNetC04_2,
            matchResultList$rseOurs04LW6_2
          ),
          title = "Effect of Manufacturing Job Loss from 2004 to 2015",
          dep.var.labels = "Change in Dem. Share, 2016-2012",
          omit.stat = c("rsq","ser"),
          keep = c("Job Losses", "White Job Losses", "Nonwhite Job Losses"),
          add.lines = list(
            c("White Share/Svc Layoffs", "No", "Yes", "Yes", "No", "Yes", "Yes"),
            c("Covariate Balancing", "No", "No", "Yes", "No", "No", "Yes")
          ),
          digits = 2,
          header = F,
          type = "latex",
          label = "regResult04", column.sep.width = "2pt", font.size = "small"
)

@@

\FloatBarrier
\begin{figure} 
\caption{Counterfactual Scenario: Change in Democratic Vote Share Assuming 25th Percentile of Manufacturing Job Losses}
\label{counterfactual1}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}

@@

\FloatBarrier
\begin{figure}

\caption{Counterfactual Simulation Results: Democratic Electoral College Votes Assuming White Manufacturing Job Losses Between 1st and 99th Percentile }
\label{counterfactual2}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}

@@

\FloatBarrier
\begin{figure} 
\caption{Counterfactual Simulation Results: Democratic Presidential Win Frequency}
\label{counterfactual3}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center", regfloat=FALSE}

@@

Restricting our analysis to only the 2012-2015 period, while continuing to use the net job loss metric, produces an effect opposite to that of \cite{Baccini21}. The loss of Democratic vote share seems to be associated with a gain in manufacturing jobs, both in general and among whites. We believe this time-restricted analysis suffers from an omitted variable bias. As discussed above, the 2012-2015 period can be broadly characterized as a slight rebound from the manufacturing job losses in the 2008 financial crisis. The counties that experience job losses from 2004-2012 are more likely to have job gains in 2012-2015 (see \ref{fig:vShapePlot1}). An apparent relationship between 2012-2015 job gains and loss of Democratic vote share in fact represents 2004-2012 job losses that may lead to both the later job gains and the loss of vote share. These results can be seen in Table \ref{regResult12}.
\FloatBarrier
```{r,  echo=F, message = F, results = "asis", regfloat=FALSE}

@@

\begin{figure} 
\caption{V-Shaped Recovery in Manufacturing Job Losses: Manufacturing Job Losses (2004-2011) by (2004-2011) Decile Groupings}
\label{fig:vShapePlot1}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center"}

@@

\begin{figure} 
\caption{V-Shaped Recovery in Manufacturing Job Losses: Manufacturing Job Losses (2012-2015) by 2004-2011 Decile Groupings}
\label{vShapePlot2}
```{r, message = F, fig.height = 2.5, fig.width = 5.5, fig.align="center"}

@@

\begin{figure} 
\caption{Covariate Balance in Aggregate}
\label{balPlot}
```{r, message = F, fig.width = 5.5, fig.align="center"}
cbpsObjs <- readRDS("../cbpsObjects.rds")
plot(cbpsObjs$fit04)

@@

\begin{figure} 
\caption{Covariate Balance in Whites}
\label{balPlotWhite}
```{r, message = F, fig.width = 5.5, fig.align="center"}
plot(cbpsObjs$fit04_white)

### 2021-12-12 Final Draft/Final-Draft.tex
figure fix, small changes

\FloatBarrier
\begin{figure} \label{ManuMap}
\caption{Concentration of Deindustrialization (2004-2015) in the eastern United States }

@@

\FloatBarrier
\begin{figure} \label{natlPlotTS}
\caption{Nationwide Change in Manufacturing Employment, 1995-2019}

@@

\begin{table}[!htbp] \centering 
  \caption{Effect of Manufacturing Job Loss from 2004 to 2015} 
  \label{regResult04} 
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
\cline{2-7} 
\\[-1.8ex] & \multicolumn{6}{c}{Change in Dem. Share, 2016-2012} \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
\hline \\[-1.8ex] 
 Job Losses & $-$0.34$^{**}$ & $-$0.38$^{**}$ & $-$0.03$^{***}$ &  &  &  \\ 
  & (0.14) & (0.17) & (0.01) &  &  &  \\ 
  & & & & & & \\ 
 White Job Losses &  &  &  & $-$0.23$^{***}$ & $-$0.32$^{***}$ & $-$0.07$^{***}$ \\ 
  &  &  &  & (0.08) & (0.10) & (0.02) \\ 
  & & & & & & \\ 
 Nonwhite Job Losses &  &  &  & 0.15$^{***}$ & 0.20$^{***}$ & 0.09$^{**}$ \\ 
  &  &  &  & (0.05) & (0.06) & (0.04) \\ 
  & & & & & & \\ 
\hline \\[-1.8ex] 
White Share/Svc Layoffs & No & Yes & Yes & No & Yes & Yes \\ 
Covariate Balancing & No & No & Yes & No & No & Yes \\ 
Observations & 3,049 & 3,049 & 2,750 & 2,750 & 2,750 & 2,750 \\ 
Adjusted R$^{2}$ & 0.72 & 0.72 & 0.74 & 0.75 & 0.75 & 0.73 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table} 
\FloatBarrier

@@

\FloatBarrier
\begin{figure} \label{counterfactual1}
\caption{Counterfactual Scenario: Change in Democratic Vote Share Assuming 25th Percentile of Manufacturing Job Losses}

@@

\FloatBarrier
\begin{figure} \label{counterfactual2}

\caption{Counterfactual Simulation Results: Democratic Electoral College Votes Assuming White Manufacturing Job Losses Between 1st and 99th Percentile }

@@

It is important to note that, in the counterfactual and simulation
exercises shown in Figures \ref{counterfactual1}, \ref{counterfactual2},
and \ref{counterfactual3}, we estimate the change in Democratic vote
share assuming various levels of white manufacturing job losses
\emph{holding all else equal, including non-white manufacturing job
losses.} Given that white and non-white manufacturing job losses are
weakly-to-moderately correlated across counties (r = .33), this is
likely a poor assumption, because, in reality, a decline in white
manufacturing job losses would likely correspond to a decline in
non-white manufacturing job losses. Given that non-white manufacturing
job losses were expected to \emph{increase} Democratic vote share in
2016, this means that our simulation results may be overestimating
effect sizes because it does not take into account this second-order
impact.

\FloatBarrier
\begin{figure} \label{counterfactual3}
\caption{Counterfactual Simulation Results: Democratic Presidential Win Frequency}

@@

Finally, we would like to acknowledge our focus on white manufacturing
job losses over non-white manufacturing job losses. In this analysis, we
focus on white manufacturing job losses for two reasons. First, in the
United States, manufacturing employment and job losses have historically
been predominately white (See Table \ref{tab:descTable}). Second, our
results suggest that the white manufacturing job losses have a larger
negative effect on Democratic vote share relative to the magnitude of
the positive effect from non-white manufacturing losses. For these
reasons, our results suggest that, in aggregate, white-manufacturing
losses have had a more significant effect on political outcomes than
non-white manufacturing job losses. However, we would like to highlight
that the political consequences from non-white deindustrialization is an
important area for future work.

\subsubsection{Timing}

Restricting our analysis to only the 2012-2015 period, while continuing
to use the net job loss metric, produces an effect opposite to that of
\cite{Baccini21}. The loss of Democratic vote share seems to be
associated with a gain in manufacturing jobs, both in general and among
whites. We believe this time-restricted analysis suffers from an omitted
variable bias. As discussed above, the 2012-2015 period can be broadly
characterized as a slight rebound from the manufacturing job losses in
the 2008 financial crisis. The counties that experience job losses from
2004-2012 are more likely to have job gains in 2012-2015 (see
\ref{vShapePlot}). An apparent relationship between 2012-2015 job gains
and loss of Democratic vote share in fact represents 2004-2012 job
losses that may lead to both the later job gains and the loss of vote
share. These results can be seen in Table \ref{regResult12}.
\FloatBarrier

@@

\begin{figure} \label{vShapePlot1}
\caption{V-Shaped Recovery in Manufacturing Job Losses: Manufacturing Job Losses (2004-2011) by (2004-2011) Decile Groupings}

@@

\FloatBarrier
\begin{figurenotes}
Deciles are determined based on our preferred measure of deindustrialization: total net manufacturing job losses from 2004-2015 divided by total 2004 employment. Counties in decile 1 had the lowest manufacturing job losses (i.e., an increase in manufacturing employment) while counties in decile 10 had the highest manufacturing job losses. The y-axis shows the mean measure of manufacturing job losses (2004-2015) per 2004 worker across counties in each decile group.
\end{figurenotes}
\end{figure}

\begin{figure} \label{vShapePlot2}
\caption{V-Shaped Recovery in Manufacturing Job Losses: Manufacturing Job Losses (2012-2015) by 2004-2011 Decile Groupings}

@@

\begin{figure} \label{balPlot}
\caption{Covariate Balance in Aggregate}

@@

\begin{figure} \label{balPlotWhite}
\caption{Covariate Balance in Whites}

>>>>

\FloatBarrier
\begin{figure} 
\caption{Concentration of Deindustrialization (2004-2015) in the eastern United States }
\label{ManuMap}

@@

\FloatBarrier
\begin{figure}
\caption{Nationwide Change in Manufacturing Employment, 1995-2019}
\label{natlPlotTS}

@@

\begin{table}[!htbp] \centering 
  \caption{Effect of Manufacturing Job Loss from 2004 to 2015} 
  \label{regResult04} 
\small 
\begin{tabular}{@{\extracolsep{2pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
\cline{2-7} 
\\[-1.8ex] & \multicolumn{6}{c}{Change in Dem. Share, 2016-2012} \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
\hline \\[-1.8ex] 
 Job Losses & $-$0.34$^{**}$ & $-$0.38$^{**}$ & $-$0.03$^{***}$ &  &  &  \\ 
  & (0.14) & (0.17) & (0.01) &  &  &  \\ 
  & & & & & & \\ 
 White Job Losses &  &  &  & $-$0.23$^{***}$ & $-$0.32$^{***}$ & $-$0.07$^{***}$ \\ 
  &  &  &  & (0.08) & (0.10) & (0.02) \\ 
  & & & & & & \\ 
 Nonwhite Job Losses &  &  &  & 0.15$^{***}$ & 0.20$^{***}$ & 0.09$^{**}$ \\ 
  &  &  &  & (0.05) & (0.06) & (0.04) \\ 
  & & & & & & \\ 
\hline \\[-1.8ex] 
White Share/Svc Layoffs & No & Yes & Yes & No & Yes & Yes \\ 
Covariate Balancing & No & No & Yes & No & No & Yes \\ 
Observations & 3,049 & 3,049 & 2,750 & 2,750 & 2,750 & 2,750 \\ 
Adjusted R$^{2}$ & 0.72 & 0.72 & 0.74 & 0.75 & 0.75 & 0.73 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table} 
\FloatBarrier

@@

\FloatBarrier
\begin{figure} 
\caption{Counterfactual Scenario: Change in Democratic Vote Share Assuming 25th Percentile of Manufacturing Job Losses}
\label{counterfactual1}

@@

\FloatBarrier
\begin{figure}

\caption{Counterfactual Simulation Results: Democratic Electoral College Votes Assuming White Manufacturing Job Losses Between 1st and 99th Percentile }
\label{counterfactual2}

@@

Finally, it is important to note that, in the counterfactual and
simulation exercises shown in Figures \ref{counterfactual1},
\ref{counterfactual2}, and \ref{counterfactual3}, we estimate the change
in Democratic vote share assuming various levels of white manufacturing
job losses \emph{holding all else equal, including non-white
manufacturing job losses.} Given that white and non-white manufacturing
job losses are weakly-to-moderately correlated across counties (r =
.33), this is likely a poor assumption, because, in reality, a decline
in white manufacturing job losses would likely correspond to a decline
in non-white manufacturing job losses. Given that non-white
manufacturing job losses were expected to \emph{increase} Democratic
vote share in 2016, this means that our simulation results may be
overestimating effect sizes because it does not take into account this
second-order impact.

\FloatBarrier
\begin{figure} 
\caption{Counterfactual Simulation Results: Democratic Presidential Win Frequency}
\label{counterfactual3}

@@

\FloatBarrier
\end{figure}
\FloatBarrier
\subsubsection{Timing}

Restricting our analysis to only the 2012-2015 period, while continuing
to use the net job loss metric, produces an effect opposite to that of
\cite{Baccini21}. The loss of Democratic vote share seems to be
associated with a gain in manufacturing jobs, both in general and among
whites. We believe this time-restricted analysis suffers from an omitted
variable bias. As discussed above, the 2012-2015 period can be broadly
characterized as a slight rebound from the manufacturing job losses in
the 2008 financial crisis. The counties that experience job losses from
2004-2012 are more likely to have job gains in 2012-2015 (see
\ref{fig:vShapePlot1}). An apparent relationship between 2012-2015 job
gains and loss of Democratic vote share in fact represents 2004-2012 job
losses that may lead to both the later job gains and the loss of vote
share. These results can be seen in Table \ref{regResult12}.
\FloatBarrier

@@

\begin{figure} 
\caption{V-Shaped Recovery in Manufacturing Job Losses: Manufacturing Job Losses (2004-2011) by (2004-2011) Decile Groupings}
\label{fig:vShapePlot1}

@@

\FloatBarrier
\begin{figurenotes}
Deciles are determined based on our preferred measure of deindustrialization: total net manufacturing job losses from 2004-2015 divided by total 2004 employment. Counties in decile 1 had the lowest manufacturing job losses (i.e., an increase in manufacturing employment) while counties in decile 10 had the highest manufacturing job losses. The y-axis shows the mean measure of manufacturing job lossess (2004-2015) per 2004 worker across counties in each decile group.
\end{figurenotes}
\end{figure}

\begin{figure} 
\caption{V-Shaped Recovery in Manufacturing Job Losses: Manufacturing Job Losses (2012-2015) by 2004-2011 Decile Groupings}
\label{vShapePlot2}

@@

\begin{figure} 
\caption{Covariate Balance in Aggregate}
\label{balPlot}

@@

\begin{figure} 
\caption{Covariate Balance in Whites}
\label{balPlotWhite}
